---
title: "TFM_Results"
author: "Jon Ortiz Abalia"
date: "2 de mayo de 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library("AnnotationDbi")
library("org.Hs.eg.db")
library(Biobase)
library(dplyr)
library(openxlsx)
library(knitr) # for kable()
library(kableExtra)
```

```{r}
# Establezco el directorio de trabajo

setwd("C:/Users/Usuario/Desktop/UOC/TFM/carles_barcelo/RStudio/GSEA")

# Cargo el archivo del dataset

datos<-read_excel("RESULTS.xlsx", sheet=3, col_types = c("text", "text","text","text","text","numeric"))

head(datos)
```

# Annotation of gene names. Addition of columns 'Entrez_ID' and 'Gene_Name' 

```{r}

GeneCol<- as.character(datos$GENE)

datos$EntrezID<-mapIds(org.Hs.eg.db, GeneCol, "ENTREZID", "SYMBOL")

datos$Gene_Name<-mapIds(org.Hs.eg.db, GeneCol, "GENENAME", "SYMBOL")


```

# Creation of a new column called 'COMMON_GENE' which contains information on whether a gene is deregulated (or not) in both cancers: 'Yes' or 'No' 


```{r}

datos<-datos %>% 
  group_by(GENE) %>%
  mutate(COMMON_GENE = case_when(any(CANCER_TYPE=='BC') & any(CANCER_TYPE=='MCL') ~ "YES", TRUE ~ "NO"))

# List of unique genes shared by BC and MCL

common_genes<-distinct(datos[which(datos$COMMON_GENE=='YES'),], GENE)


```

# Creation of new columns with the information of in how many datasets, in how many BC datasets and in how many MCL datasets the common gene has been found deregulated


```{r}

## TOTAL DATASETS

# Create table with each gene and number of total datasets in which the gene has been found deregulated

table1<-datos %>% 
  filter(COMMON_GENE=='YES') %>% 
  group_by(GENE) %>%
  tally()

names(table1)[2]<-'NUM_DATASETS'

# Create the corresponding column with the information of total datasets

datos2<-datos %>% 
  filter(COMMON_GENE=='YES') %>% 
  group_by(GENE) %>%
  mutate(DATASETS_TOT= n())

## BC DATASETS

# Create table with each gene and number of BC datasets in which the gene has been found deregulated

table2<-datos %>% 
  filter(COMMON_GENE=='YES' & CANCER_TYPE=='BC') %>% 
  group_by(GENE) %>%
  tally()

names(table2)[2]<-'BC_DATASETS'


## MCL DATASETS

# Create table with each gene and number of BC datasets in which the gene has been found deregulated

table3<-datos %>% 
  filter(COMMON_GENE=='YES' & CANCER_TYPE=='MCL') %>% 
  group_by(GENE) %>%
  tally()

names(table3)[2]<-'MCL_DATASETS'

# Join tables

datos2<-merge(datos2, table2, by="GENE",all.x=TRUE)
datos2<-merge(datos2, table3, by="GENE",all.x=TRUE)

```

# New columns to record the informarion of the number of datasets in which the gene has been upregulated, downregulated in both cancer types and in total


```{r}

# Number of total datasets in which the gene has been found upregulated

table4<-datos2 %>% 
  filter(`UP/DOWNREGULATED`=='UP') %>%
  group_by(GENE) %>%
  tally()

names(table4)[2]<-'TOT_UPREGULATED'

# Number of total datasets in which the gene has been found downpregulated

table5<-datos2 %>% 
  filter(`UP/DOWNREGULATED`=='DOWN') %>%
  group_by(GENE) %>%
  tally()

names(table5)[2]<-'TOT_DOWNREGULATED'

# Number of BC datasets in which the gene has been found UPpregulated

table6<-datos2 %>% 
  filter(`UP/DOWNREGULATED`=='UP' & CANCER_TYPE=='BC') %>%
  group_by(GENE) %>%
  tally()

names(table6)[2]<-'BC_UPREGULATED'

# Number of BC datasets in which the gene has been found downpregulated

table7<-datos2 %>% 
  filter(`UP/DOWNREGULATED`=='DOWN' & CANCER_TYPE=='BC') %>%
  group_by(GENE) %>%
  tally()

names(table7)[2]<-'BC_DOWNREGULATED'

# Number of MCL datasets in which the gene has been found UPpregulated

table8<-datos2 %>% 
  filter(`UP/DOWNREGULATED`=='UP' & CANCER_TYPE=='MCL') %>%
  group_by(GENE) %>%
  tally()

names(table8)[2]<-'MCL_UPREGULATED'

# Number of MCL datasets in which the gene has been found downpregulated

table9<-datos2 %>% 
  filter(`UP/DOWNREGULATED`=='DOWN' & CANCER_TYPE=='MCL') %>%
  group_by(GENE) %>%
  tally()

names(table9)[2]<-'MCL_DOWNREGULATED'

# Join tables 


datos2<-merge(datos2, table4, by="GENE",all.x=TRUE)
datos2<-merge(datos2, table5, by="GENE",all.x=TRUE)
datos2<-merge(datos2, table6, by="GENE",all.x=TRUE)
datos2<-merge(datos2, table7, by="GENE",all.x=TRUE)
datos2<-merge(datos2, table8, by="GENE",all.x=TRUE)
datos2<-merge(datos2, table9, by="GENE",all.x=TRUE)

# Replace NA values by '0'

datos2$TOT_UPREGULATED[which(is.na(datos2$TOT_UPREGULATED))]<-0
datos2$TOT_DOWNREGULATED[which(is.na(datos2$TOT_DOWNREGULATED))]<-0
datos2$BC_UPREGULATED[which(is.na(datos2$BC_UPREGULATED))]<-0
datos2$BC_DOWNREGULATED[which(is.na(datos2$BC_DOWNREGULATED))]<-0
datos2$MCL_UPREGULATED[which(is.na(datos2$MCL_UPREGULATED))]<-0
datos2$MCL_DOWNREGULATED[which(is.na(datos2$MCL_DOWNREGULATED))]<-0


```

# Analyze gene sets


```{r}

datos3<-distinct(datos2, GENE, GENE_SET)

table(datos3$GENE_SET)

```

# Createnew file deleting columns

\newpage

```{r echo=FALSE}

datos_def<-distinct(datos2, GENE, .keep_all=TRUE)
datos_def<-datos_def[,-c(2,3,4,5,9)]

table<-datos_def[,c(1,3,4)]

kable(table,format='latex', booktabs=T, align='lll', longtable=TRUE) %>%
  row_spec(0, bold=TRUE) %>%
  kable_styling(full_width=F, latex_options = c("hold_position", "repeat_header"))


```

# Exports

```{r}
write.xlsx(datos2, file = "gene_Signature.xlsx",sheetName = "annot", append = FALSE)

write.xlsx(datos, file = "gsea_results.xlsx",sheetName = "annot", append = FALSE)

write.xlsx(datos_def, file = "common_genes.xlsx",sheetName = "annot", append = FALSE)


```




